---
title: '6:W43: Practicing functions with Gapminder'
author: "Louise Brix Pilegaard Hansen"
date: '2022-10-27'
output: html_document
---



### Loading the data and packages
```{r message=FALSE, warning=FALSE}
library(gapminder)
pacman::p_load("tidyverse")
pacman::p_load("dplyr")
df <- as.data.frame(gapminder)
```


# 1)

Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. You can use the population and GDPpercapita columns for it. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.


```{r}
gdp_fun <- function(country, year) {
  stopifnot(is.character(country)) # stop function if country is not a character
  stopifnot(is.numeric(year)) # stop function if year is not numeric
  
  row <- which(df$country == country & df$year == year) # finding the row of the data with the correct country and year 
  gdppercap <- df[row, 6] # extracting the gdppercap (column 6)
  pop <- df[row, 5] # extracting the population size (column 5)
  gdp <- gdppercap*pop # calculating gdp
  return(gdp)
}
```

## calculating GDP for Denmark 
```{r}
years <- c(1967, 1977, 1987, 1997, 2007, 2017)

for (year in years){
  if (year %in% df$year){ # if the year is in the gapminder dataset, continue
    gdp <- gdp_fun("Denmark", year) # running the function
    print(paste0("In ", year, " the GDP for Denmark was ", gdp)) # preparing output
  }
  else {
    print(paste0("No data for ", year))
  }
}
```


# 2)

Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and prints out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70. (Hint: remember the grepl function, and review the Control Flow tutorial)

```{r}
for (i in unique(df$country)){ # for every unique country in the dataset

  if (grepl('^B', i)){ # if country starts with 'B'
  
  country_dat <- df %>% # creating dataframe with the country's data
    filter(country == i)
  
  lifeexp_mean <- mean(country_dat$lifeExp) # calculating mean life expectancy for each country
  
  if (lifeexp_mean < 50) {
    print(paste0(i, " has a mean life expectancy of ", round(lifeexp_mean), " years, i.e., below 50 years"))
  }
  
  else if (50 < lifeexp_mean & lifeexp_mean < 70){
    print(paste0(i, " has a mean life expectancy of ", round(lifeexp_mean), " years, i.e., between 50 and 70 years"))
  } 
  else if (lifeexp_mean > 70){
    print(paste0(i, " has a mean life expectancy of ", round(lifeexp_mean), " years, i.e., over 70 years"))
  }
  }
}
```
