---
title: "billboard_scrape"
author: "Louise Brix Pilegaard Hansen"
date: '2022-12-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#install.packages('rvest')
library(rvest)
library(tidyverse)
library(geniusr)
```

# Scraping billboard.com

## could be improved
```{r define scraper function}
billboard_scraper <- function(url){
  url_html <- read_html(url)
  
  entries <- url_html %>% 
    html_elements("li")
  
  titles <- entries %>% 
    html_elements(".c-title") %>% 
    html_text2() 
  
  artists <- url_html %>%
    html_nodes(".c-label.a-no-trucate") %>%
    html_text2()

  return(data.frame(title = titles, artist = artists))
}
```

# could be improved (don't run )
```{r define URLs}
# pick weeks and years 
urls_1982 <- c("https://www.billboard.com/charts/hot-100/1982-01-09/", "https://www.billboard.com/charts/hot-100/1982-04-10/", "https://www.billboard.com/charts/hot-100/1982-07-10/", "https://www.billboard.com/charts/hot-100/1982-10-09/")

urls_2002 <- c("https://www.billboard.com/charts/hot-100/2002-01-12/", "https://www.billboard.com/charts/hot-100/2002-04-13/", "https://www.billboard.com/charts/hot-100/2002-07-13/", "https://www.billboard.com/charts/hot-100/2002-10-12/")

urls_2022 <- c("https://www.billboard.com/charts/hot-100/2022-01-08/", "https://www.billboard.com/charts/hot-100/2022-04-09/", "https://www.billboard.com/charts/hot-100/2022-07-09/", "https://www.billboard.com/charts/hot-100/2022-10-08/")
```


# could be improved (don't run)
```{r scrape billboard URLs}
#url_vec <- c(urls_1982, urls_2002, urls_2022)

playlist_1982 <- lapply(urls_1982, billboard_scraper)
playlist_1982 <- do.call(rbind, playlist_1982)
playlist_1982$year <- "1982"
write_csv(playlist_1982, "playlist_1982.csv")

playlist_2002 <- lapply(urls_2002, billboard_scraper)
playlist_2002 <- do.call(rbind, playlist_2002)
playlist_2002$year <- "2002"
write_csv(playlist_2002, "playlist_2002.csv")

playlist_2022 <- lapply(urls_2022, billboard_scraper)
playlist_2022 <- do.call(rbind, playlist_2022)
playlist_2022$year <- "2022"
write_csv(playlist_2022, "playlist_2022.csv")
```


# Getting lyrics from genius.com using the Genius API 

I have created an environment variable containing my personal API token to genius.com. This key cannot be shared; to create your own, go to https://genius.com/api-clients

## read in all the songs
```{r reading in data dont run}
playlist_1982 <- read_csv("playlist_1982.csv")
playlist_2002 <- read_csv("playlist_2002.csv")
playlist_2022 <- read_csv("playlist_2022.csv")

#playlist_1982$title[15] <- "Why Do Fools Fall In Love"  

playlists <- list(playlist_1982, playlist_2002, playlist_2022)

# replace lil nas x manually
# fjern alt efter featuring

playlist_clean <- function(df){
  df$artist <- str_replace_all(df$artist, "&", "and")
  df$artist <- str_replace_all(df$artist, "(\\s[xX]\\s)", " and ")
  df$artist <- str_replace_all(df$artist, ",", " and")
  df$artist <- str_replace_all(df$artist, "((Featuring))(?s)(.*$)", "")
  return(df)
}

playlists <- lapply(playlists, playlist_clean)

playlists[[3]]$artist <- str_replace_all(playlists[[3]]$artist, "Lil Nas and", "Lil Nas X")
```

```{r delete}
song <- search_genius(search_term = "What Happened To Virgil Lil Durk")#search_genius(search_term = playlist$title[8])


id <- song$content[[1]]$id

get_lyrics_id(song_id = id)

hej <- search_genius(paste0("Why Do Fools Fall In Love", " - ", "Diana Ross"))
```


```{r creating tryout-subset}
#try82 <- playlist_1982[1:10, ]
#try02 <- playlist_2002[1:10, ]
#try22 <- playlist_2022[1:10, ]
#playlists_try <- list(try82, try02, try22)
```


```{r activate api token}
Sys.getenv('GENIUS_API_TOKEN') # getting the value of the API token
genius_token() # activating it
```

```{r}
## replace all X + & with "and". 
## check that songs name match rather than song_id
## remove songs that can't be found
```


### add ID and language
```{r dont run}
add_id <- function(playlist){
  for (i in 1:length(playlist$title)){
  song <- playlist$title[i]
  artist <- playlist$artist[i]
  print(song)
  g_search <- try(search_genius(paste0(song, " - ", artist)))
  playlist$id[i] <- try(g_search$content[[1]]$id)
  try(playlist$lang[i] <- g_search$content[[1]]$language)
  }
  
  #playlist <- playlist %>% 
  #filter(lang == "en")
  
  return(playlist)
}
```


```{r dont run}
playlists <- lapply(playlists, add_id)

# saving to csv to be sure
all_songs <- do.call(rbind, playlists)
all_songs <- all_songs %>% 
  filter(lang == "en") %>% 
  filter(!str_detect(id, 'Error'))

write.csv(all_songs, "all_songs.csv")
```

## getting the lyrics
```{r}
lyrics_scraper <- function(df){
  lyrics_df <- NULL # initializing empty df
  for (i in 1:length(df$id)){ # looping over each song ID
    id <- df$id[i]
    lyrics <- try((get_lyrics_id(song_id = id)), silent = FALSE) # scraping lyrics
    lyrics_df <- rbind(lyrics_df, lyrics)
  }
 return(lyrics_df)
}
```


```{r getting lyrics tryout}
#playlist_lyrics_tryout <- lapply(playlists_try, lyrics_scraper) # returns a list of three dataframes, one for each year
```

```{r reading in playlist df}
playlists_all <- read_csv("all_songs.csv")

p_1982 <- playlists_all %>% 
  filter(year == 1982) %>% 
  distinct(id, .keep_all = TRUE)

p_2002 <- playlists_all %>% 
  filter(year == 2002) %>% 
  distinct(id, .keep_all = TRUE)

p_2022 <- playlists_all %>% 
  filter(year == 2022) %>% 
  distinct(id, .keep_all = TRUE)

playlists <- list(p_1982, p_2002, p_2022)
```


```{r getting lyrics}
lyrics <- lapply(playlists, lyrics_scraper)

write.csv(lyrics[[1]], "lyrics_1982.csv")
write.csv(lyrics[[2]], "lyrics_2002.csv")
write.csv(lyrics[[3]], "lyrics_2022.csv")
```

```{r}
#lyrics_full_df <- do.call(rbind, lyrics)

write.csv(lyrics_full_df, "lyrics_full_df.csv")

write.csv(lyrics[[1]], "lyrics_1982.csv")
write.csv(lyrics[[2]], "lyrics_2002.csv")
write.csv(lyrics[[3]], "lyrics_2022.csv")
```

```{r}
lyrics_1982 <- read_csv("lyrics_1982.csv")
lyrics_2002 <- read_csv("lyrics_2002.csv")
lyrics_2022 <- read_csv("lyrics_2022.csv")

lyrics <- list(lyrics_1982, lyrics_2002, lyrics_2022)
```


den importerer nogle mærkelige sange ved dem, den fejler på - bør måske bare slette dem i stedet?

The get_lyrics_id function somehow fails if you run it on many songs (I have tried searching for the song-ids that fail manually, and it works, so it's maybe the size of the playlist that makes it fail). Thus, the df requires some cleaning and running the function again on the missing songs: 

```{r}
for (i in 1:length(lyrics)){
  #playlist <- playlist_lyrics_tryout[[i]]
  ind <- which(playlists[[i]]$id %in% lyrics[[i]]$song_id)
  songs_missing <- playlists[[i]]$title[-ind]
  print(paste0("In playlist no ", i, " ", songs_missing, " is missing"))
  
  for (song in songs_missing){
    id <- playlists[[i]]$id[which(playlists[[i]]$title == song)]
    lyrics_df <- try((get_lyrics_id(song_id = id)), silent = FALSE)
    lyrics[[i]] <- rbind(lyrics[[i]], lyrics_df)
  }
  lyrics[[i]] <- lyrics[[i]] %>% 
    filter(!str_detect(song_id, 'Error'))
  
  print(lyrics[[i]])
}
```

```{r}
ind <- which(playlists[[3]]$artist %in% lyrics[[3]]$artist_name)
songs_missing <- playlists[[3]]$title[-ind]
songs_missing
```

```{r}
tryout123 <- lyrics[[3]]
```


# remove NA's, certain songs and errors


```{r}
cleaner_fun <- function(df){
  df <- df %>% 
  filter(artist_name != "Spotify") %>% 
  filter(!str_detect(song_id, 'Error')) %>% 
  filter(!str_detect(artist_name, 'Genius')) %>% 
  drop_na()
  
  return(df)
}

lyrics <- lapply(lyrics, cleaner_fun)
```

```{r}
remove_82 <- c("6868363","4033274","480443","748742","581048","467863","491080","387797","297338","1868856","504292","2644622","5461894","2831404")

remove_02 <- c("2422119", "3299443", "8176969", "409534", "3038244", "3038244")

remove_22 <- c("7559954", "7507324", "7457331", "8687863", "8631841", "8451026", "7130381")

#remove <- list(remove_82, remove_02, remove_22)

```

```{r}
lyrics[[1]] <- lyrics[[1]] %>% 
  filter(! song_id %in% remove_82)

lyrics[[2]] <- lyrics[[2]] %>% 
  filter(! song_id %in% remove_02)

lyrics[[3]] <- lyrics[[3]] %>% 
  filter(! song_id %in% remove_22)
```

```{r}
write.csv(lyrics[[1]], "lyrics_1982.csv")
write.csv(lyrics[[2]], "lyrics_2002.csv")
write.csv(lyrics[[3]], "lyrics_2022.csv")
```

